---
title: "First Try"
author: 
- "Yoni Slutzky"
- "Sagi Ahrac"
date: "12/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read Data
```{r}
Train_Features <- read.table("./data/train_ratings_all.dat", header = F, sep = "\t")
Train_Dates_Features <- read.table("./data/train_dates_all.dat", header = F, sep = "\t")
Train_Labels <- read.table("./data/train_y_rating.dat", header = F, sep = "\t")
Train_Dates_Labels <- read.table("./data/train_y_date.dat", header = F, sep = "\t")

Test_Features <- read.table("./data/test_ratings_all.dat", header = F, sep = "\t")
Test_Dates_Features <- read.table("./data/test_dates_all.dat", header = F, sep = "\t")
Test_Dates_Labels <- read.table("./data/test_y_date.dat", header = F, sep = "\t")

movie_info <- read.table("./data/movie_titles.txt", header = F, sep = ',')
```

Name the Data
```{r}
colnames(movie_info) <- c("Release Years", "Names", "Genres")
colnames(Train_Features) <- movie_info$Names
colnames(Test_Features) <- colnames(Train_Features)
colnames(Train_Labels) <- c('y')
```

Train processing - Replace zeros by average rating of each user in the genre
```{r}
non_zero_avg = rep(0, 10000)
for (i in 1:10000) {
  non_zero_avg[i] = mean(unlist(Train_Features[i, ])[as.vector(Train_Features[i, ] != 0)])
}
  
avg_per_genre <- data.frame(matrix(nrow = 10000, ncol = 7))
colnames(avg_per_genre) = unique(movie_info$Genres)
for (genre in unique(movie_info$Genres)){
  avg_per_genre[genre] <- rowSums(Train_Features[, movie_info$Genres == genre]) / 
    rowSums(!!Train_Features[, movie_info$Genres == genre])
}
for (i in 1:10000) {
  for (j in 1:99) {
    if (Train_Features[i, j] == 0) {
      Train_Features[i,j] = avg_per_genre[i, movie_info$Genres[j]]
      if (is.na(Train_Features[i, j])){
        Train_Features[i,j] = non_zero_avg[i]
      }
    }
  }
}
```


Test processing - Replace zeros by average rating of each user in the genre (to be used for prediction)
```{r}
non_zero_avg = rep(0, 2931)
for (i in 1:2931) {
  non_zero_avg[i] = mean(unlist(Test_Features[i, ])[as.vector(Test_Features[i, ] != 0)])
}
  
avg_per_genre <- data.frame(matrix(nrow = 2931, ncol = 7))
colnames(avg_per_genre) = unique(movie_info$Genres)
for (genre in unique(movie_info$Genres)){
  avg_per_genre[genre] <- rowSums(Test_Features[, movie_info$Genres == genre]) / 
    rowSums(!!Test_Features[, movie_info$Genres == genre])
}
for (i in 1:2931) {
  for (j in 1:99) {
    if (Test_Features[i, j] == 0) {
      Test_Features[i,j] = avg_per_genre[i, movie_info$Genres[j]]
      if (is.na(Test_Features[i, j])){
        Test_Features[i,j] = non_zero_avg[i]
      }
    }
  }
}
```

Create train and CV sets
```{r}
CV_subset <- sample(10000, 2000)
train_set <- data.frame(Train_Features[-CV_subset, ], y = Train_Labels[-CV_subset, ])
CV_set <- data.frame(Train_Features[CV_subset, ], y = Train_Labels[CV_subset, ])
```

Ridge Regression with Cross-validation
```{r}
library(MASS)
lambdas <- exp(seq(-15, 12, by = 0.1))
models <- lm.ridge(y~., data = train_set, lambda = lambdas)
```

Train RMSE
```{r}
preds_on_train <- as.matrix(train_set[, 1:99]) %*% t(coef(models)[, -1]) + rep(1, 8000) %o% coef(models)[, 1]
resid_on_train <- train_set[, 1] - preds_on_train
rss_on_train <- apply(resid_on_train^2, 2, sum)
plot(models$lambda, sqrt(rss_on_train/8000))
min(sqrt(rss_on_train/8000))
```

CV RMSE
```{r}
preds_on_CV <- as.matrix(CV_set[, 1:99]) %*% t(coef(models)[, -1]) + rep(1, 2000) %o% coef(models)[, 1]
resid_on_CV <- CV_set[, 1] - preds_on_CV
rss_on_CV <- apply(resid_on_CV^2, 2, sum)
plot(models$lambda, sqrt(rss_on_CV/2000))
min(sqrt(rss_on_CV/2000))
```


Lasso Regression with Cross-validation
```{r}
library(glmnet)
lambdas <- exp(seq(-15, 0, by = 0.1))
models <- glmnet(train_set[, 1:99], train_set[, 100], lambda = lambdas,
                 family = "gaussian",
                 intercept = T, alpha = 1)
```

Train RMSE
```{r}
preds_on_train <- predict(models, newx = as.matrix(train_set[, 1:99]))
resid_on_train <- apply(preds_on_train, 2, "-", train_set[, 100])
rss_on_train <- apply(resid_on_train^2, 2, sum)
plot(models$lambda, sqrt(rss_on_train / 8000))
min(sqrt(rss_on_train / 8000))
```

CV RMSE
```{r}
preds_on_CV <- predict(models, newx = as.matrix(CV_set[, 1:99]))
resid_on_CV <- apply(preds_on_CV, 2, "-", CV_set[, 100])
rss_on_CV <- apply(resid_on_CV^2, 2, sum)
plot(models$lambda, sqrt(rss_on_CV / 2000))
min(sqrt(rss_on_CV / 2000))
```

Create complete train and test sets
```{r}
train_set <- data.frame(Train_Features, y = Train_Labels)
test_set <- data.frame(Test_Features)
```

Retrain on all of training data using Lasso
```{r}
lambdas <- exp(seq(-15, 0, by = 0.1))
models <- glmnet(train_set[, 1:99], train_set[, 100], lambda = lambdas,
                 family = "gaussian",
                 intercept = T, alpha = 1)
```

Train RMSE
```{r}
preds_on_train <- predict(models, newx = as.matrix(train_set[, 1:99]))
resid_on_train <- apply(preds_on_train, 2, "-", train_set[, 100])
rss_on_train <- apply(resid_on_train^2, 2, sum)
rmse_on_train <- sqrt(rss_on_train / 10000)
plot(models$lambda, rmse_on_train)
min(rmse_on_train)
```

Predict on test
```{r}
minimal_index <- 1
minimal_rmse <- 1
for (i in 1:length(lambdas)) {
  if (rmse_on_train[i] <= minimal_rmse){
    minimal_index <- i
    minimal_rmse <- rmse_on_train[i]
  }
}
preds_on_test <- predict(models, newx = as.matrix(test_set[, 1:99]))[, i]
write.csv(preds_on_test, file = "./attempts/Second_Try.csv")
```







