---
title: "Third Try"
author: 
- "Yoni Slutzky"
- "Sagi Ahrac"
date: "12/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read Data
```{r}
tr <- read.table("./data/train_ratings_all.dat", header = F, sep = "\t")
tr_pr <- read.csv("./data/Train_Features_Processed.csv")[, 2:100]
tr_y <- read.table("./data/train_y_rating.dat", header = F, sep = "\t")
movie_info <- read.table("./data/movie_titles.txt", header = F, sep = ',')
```


Name the Data
```{r}
colnames(movie_info) <- c("Release Years", "Names", "Genres")
colnames(tr) <- movie_info$Names
colnames(tr_pr) <- movie_info$Names
colnames(tr_y) <- c('y')
```


Train - CV split
```{r}
CV_subset <- sample(10000, 2000)

CV <- data.frame(tr[CV_subset, ])
CV_y <- tr_y[CV_subset, ]

tr <- data.frame(tr[-CV_subset, ])
tr_y <- tr_y[-CV_subset, ]

CV_pr <- data.frame(tr_pr[CV_subset, ])
tr_pr <- data.frame(tr_pr[-CV_subset, ])
```


RMSE function
```{r}
rmse <- function(true_labels, pred_labels){
  return (sqrt(mean((true_labels-pred_labels)^2)))
}
```


Baseline RMSE on linear model without filling missing data without limits
```{r}
lin_model <- lm(y~., data=data.frame(tr, y=tr_y))
baseline <- rmse(CV_y, predict(lin_model, newdata = data.frame(CV)))
baseline
```


Linear model with filling missing data without limits
```{r}
lin_model <- lm(y~., data=data.frame(tr_pr, y=tr_y))
score <- rmse(CV_y, predict(lin_model, newdata = data.frame(CV_pr)))
baseline / score
diff <- c(baseline, score)
names(diff) <- c("baseline", "score")
diff
```


Linear model without filling missing data with limits
```{r}
lin_model <- lm(y~., data=data.frame(tr, y=tr_y))
score <- rmse(CV_y, pmax(pmin(predict(lin_model, newdata = data.frame(CV)), 5), 1))
baseline / score
diff[2] <- score
diff
```


Linear model with filling missing data
```{r}
lin_model <- lm(y~., data=data.frame(tr_pr, y=tr_y))
score <- rmse(CV_y, pmax(pmin(predict(lin_model, newdata = data.frame(CV_pr)), 5), 1))
baseline / score
diff[2] <- score
diff
```


Ridge regression without filling missing data with limits
```{r}
library(MASS)
lambdas <- exp(seq(-15, 12, by = 0.1))
models <- lm.ridge(y~., data = data.frame(tr, y=tr_y), lambda = lambdas)
preds_on_CV <- pmin(pmax(as.matrix(CV) %*% t(coef(models)[, -1]) + rep(1, length(CV_y)) %o% coef(models)[, 1], 1), 5)
rmse_on_CV <- apply(preds_on_CV, 2, rmse, true_labels = CV_y)
plot(models$lambda, rmse_on_CV)
score <- min(rmse_on_CV)
baseline / score
diff[2] <- score
diff
```


Ridge regression with filling missing data with limits
```{r}
models <- lm.ridge(y~., data = data.frame(tr_pr, y=tr_y), lambda = lambdas)
preds_on_CV <- pmin(pmax(as.matrix(CV_pr) %*% t(coef(models)[, -1]) + rep(1, length(CV_y)) %o% coef(models)[, 1], 1), 5)
rmse_on_CV <- apply(preds_on_CV, 2, rmse, true_labels = CV_y)
plot(models$lambda, rmse_on_CV)
score <- min(rmse_on_CV)
baseline / score
diff[2] <- score
diff
```


Lasso regression without filling missing data with limits
```{r}
library(glmnet)
lambdas <- exp(seq(-15, 0, by = 0.1))
models <- glmnet(tr, tr_y, lambda = lambdas,
                 family = "gaussian",
                 intercept = T, alpha = 1)
preds_on_CV <- pmin(pmax(predict(models, newx = as.matrix(CV)), 1), 5)
rmse_on_CV <- apply(preds_on_CV, 2, rmse, true_labels = CV_y)
plot(models$lambda, rmse_on_CV)
score <- min(rmse_on_CV)
baseline / score
diff[2] <- score
diff
```


Lasso regression with filling missing data with limits
```{r}
models <- glmnet(tr_pr, tr_y, lambda = lambdas,
                 family = "gaussian",
                 intercept = T, alpha = 1)
preds_on_CV <- pmin(pmax(predict(models, newx = as.matrix(CV_pr)), 1), 5)
rmse_on_CV <- apply(preds_on_CV, 2, rmse, true_labels = CV_y)
plot(models$lambda, rmse_on_CV)
score <- min(rmse_on_CV)
baseline / score
diff[2] <- score
diff
```


Elastic net without filling missing data with limits
```{r}
min_rmses <- rep(0, 100)
for (a in 1:100) {
  lambdas <- exp(seq(-15, 0, by = 0.1))
  models <- glmnet(tr, tr_y, lambda = lambdas,
                   family = "gaussian",
                   intercept = T, alpha = (1/a))
  preds_on_CV <- pmin(pmax(predict(models, newx = as.matrix(CV)), 1), 5)
  rmse_on_CV <- apply(preds_on_CV, 2, rmse, true_labels = CV_y)
  min_rmses[a] <- min(rmse_on_CV)
}
plot(seq(1, 0.01, by=-0.01), min_rmses)
score <- min(min_rmses)
baseline / score
diff[2] <- score
diff
```



Elastic net with filling missing data with limits
```{r}
min_rmses <- rep(0, 100)
for (a in 1:100) {
  lambdas <- exp(seq(-15, 0, by = 0.1))
  models <- glmnet(tr_pr, tr_y, lambda = lambdas,
                   family = "gaussian",
                   intercept = T, alpha = (1/a))
  preds_on_CV <- pmin(pmax(predict(models, newx = as.matrix(CV_pr)), 1), 5)
  rmse_on_CV <- apply(preds_on_CV, 2, rmse, true_labels = CV_y)
  min_rmses[a] <- min(rmse_on_CV)
}
plot(seq(1, 0.01, by=-0.01), min_rmses)
score <- min(min_rmses)
baseline / score
diff[2] <- score
diff
```








