---
title: "Third Try"
author: 
- "Yoni Slutzky"
- "Sagi Ahrac"
date: "12/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


BASE RMSE
Read Data
```{r}
base_tr <- read.table("./data/train_ratings_all.dat", header = F, sep = "\t")
tr_y <- read.table("./data/train_y_rating.dat", header = F, sep = "\t")
movie_info <- read.table("./data/movie_titles.txt", header = F, sep = ',')
```

Name the Data
```{r}
colnames(movie_info) <- c("Release Years", "Names", "Genres")
colnames(base_tr) <- movie_info$Names
colnames(tr_y) <- c('y')
```

Create test and cv sets
```{r}
n <- length(unlist(tr_y))
cv_subset = sample(n, 2000)
cv_base <- base_tr[cv_subset, ]
cv_y <- tr_y[cv_subset, ]
base_tr_tr <- base_tr[-cv_subset, ]
tr_tr_y <- tr_y[-cv_subset, ]
```

RMSE function
```{r}
rmse <- function(true_labels, pred_labels){
  return (sqrt(mean((true_labels-pred_labels)^2)))
}
```

RMSE on basic linear model
```{r}
model <- lm(y~., data = data.frame(X = base_tr_tr, y = tr_tr_y))
preds <- predict(model, newdata = data.frame(X = cv_base))
baseline <- rmse(cv_y, preds)
baseline
```

Read processed data
```{r}
tr <- read.csv("./data/Train_Features_Processed.csv")[, 2:100]
test <- read.csv("./data/Test_Features_Processed.csv")[, 2:100]
```


Name the Data
```{r}
colnames(tr) <- c(movie_info$Names)
colnames(test) <- c(movie_info$Names)
```


Create CV and training sets
```{r}
cv <- tr[cv_subset, ]
tr_tr <- tr[-cv_subset,]
```


RMSE on basic linear model with limits (filled missing values, added dates)
```{r}
model <- lm(y~., data = data.frame(X = tr_tr, y = tr_tr_y))
preds <- pmin(pmax(predict(model, newdata = data.frame(X = cv)), 1), 5)
score <- rmse(cv_y, preds)
(baseline / score - 1) * 100
```

RMSE on ridge regression with limits (filled missing values, added dates)
```{r}
library(MASS)
lambdas <- exp(seq(-15, 12, by = 0.1))
models <- lm.ridge(y~., data = data.frame(tr_tr, y=tr_tr_y), lambda = lambdas)
preds <- pmin(pmax(as.matrix(cv) %*% t(coef(models)[, -1]) + rep(1, length(cv_y)) %o% coef(models)[, 1], 1), 5)
rmse_on_cv <- apply(preds, 2, rmse, true_labels = cv_y)
plot(models$lambda, rmse_on_cv)
score <- min(rmse_on_cv)
baseline / score
score
```


RMSE on ridge regression with limits (filled missing values, added dates)
```{r}
library(glmnet)
lambdas <- exp(seq(-15, 0, by = 0.1))
models <- glmnet(tr_tr, tr_tr_y, lambda = lambdas,
                 family = "gaussian",
                 intercept = T, alpha = 1)
preds <- pmin(pmax(predict(models, newx = as.matrix(cv)), 1), 5)
rmse_on_cv <- apply(preds, 2, rmse, true_labels = cv_y)
plot(models$lambda, rmse_on_cv)
score <- min(rmse_on_cv)
baseline / score
score
```


KNN with correlation matrix with bias adjustment
```{r}
cor_mat <- cor(t(tr))[cv_subset, -cv_subset]
rmse_on_cv <- rep(NA, 100)
for (k in 1:100){
  preds <- rep(NA, 2000)
  for (i in 1:2000){
    nei <- order(cor_mat[i,], decreasing = T)[1:k]
    if (sum(is.na(cor_mat[i, nei])) > 0) {
      print(i)
      break
    }
    adj_y <- tr_tr_y[nei] - apply(tr_tr[nei, ], MARGIN = 1, mean)
    preds[i] <- (cor_mat[i, nei] %*% adj_y) / sum(abs(cor_mat[i, nei])) +
      mean(unlist(cv[i, ]))
  }
  rmse_on_cv[k] <- rmse(cv_y, preds)
}
(baseline / min(rmse_on_cv) - 1) * 100
```

