---
title: "First Try"
author: 
- "Yoni Slutzky"
- "Sagi Ahrac"
date: "12/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read Data
```{r}
tr <- read.table("./data/train_ratings_all.dat", header = F, sep = "\t")
tr_pr <- read.csv("./data/Train_Features_Processed.csv")[, 2:100]
tr_y <- read.table("./data/train_y_rating.dat", header = F, sep = "\t")
movie_info <- read.table("./data/movie_titles.txt", header = F, sep = ',')
```


Name the Data
```{r}
colnames(movie_info) <- c("Release Years", "Names", "Genres")
colnames(tr) <- movie_info$Names
colnames(tr_pr) <- movie_info$Names
colnames(tr_y) <- c('y')
```


PCA calculation
```{r}
tr <- data.frame(tr, tr^2, tr^3)
tr_pr <- data.frame(tr_pr, tr_pr^2, tr_pr^3)
tr_pca <- prcomp(tr,retx = T, center = T, scale = F)$x
tr_pr_pca <- prcomp(tr_pr,retx = T, center = T, scale = F)$x
```



Train - CV split
```{r}
CV_subset <- sample(10000, 2000)

CV_y <- tr_y[CV_subset, ]
tr_y <- tr_y[-CV_subset, ]

CV <- tr[CV_subset, ]
tr <- tr[-CV_subset, ]

CV_pr <- tr_pr[CV_subset, ]
tr_pr <- tr_pr[-CV_subset, ]

CV_pca <- tr_pca[CV_subset, ]
tr_pca <- tr_pca[-CV_subset, ]

CV_pr_pca <- tr_pr_pca[CV_subset, ]
tr_pr_pca <- tr_pr_pca[-CV_subset, ]
```


RMSE function
```{r}
rmse <- function(true_labels, pred_labels){
  return (sqrt(mean((true_labels-pred_labels)^2)))
}
```


Baseline RMSE on linear model without filling missing data
```{r}
lin_model <- lm(y~., data=data.frame(tr[, 1:99], y=tr_y))
baseline <- rmse(CV_y, predict(lin_model, newdata = data.frame(CV[, 1:99])))
baseline
```


RMSE on linear model with filling missing data with limits
```{r}
lin_model <- lm(y~., data=data.frame(tr_pr, y=tr_y))
score <- rmse(CV_y, pmax(pmin(predict(lin_model, newdata = data.frame(CV_pr)), 5), 1))
(baseline/score - 1) * 100
diff <- c(baseline, score)
names(diff) <- c("baseline", "score")
diff
```


SVM on PCA without filling missing data with linear kernel
```{r}
library(e1071)
model <- svm(y~., data = data.frame(tr_pca[, 1:10], y=tr_y), kernel = "linear", cost = 10, scale = F)
preds_on_CV <- as.numeric(predict(model, newdata = CV_pca[, 1:10]))
score <- rmse(CV_y, preds_on_CV)
(baseline/score - 1) * 100
diff <- c(baseline, score)
names(diff) <- c("baseline", "score")
diff
```


SVM on PCA with filling missing data with linear kernel
```{r}
model <- svm(y~., data = data.frame(tr_pr_pca[, 1:10], y=tr_y), kernel = "linear", cost = 10, scale = F)
preds_on_CV <- as.numeric(predict(model, newdata = CV_pr_pca[, 1:10]))
score <- rmse(CV_y, preds_on_CV)
(baseline/score - 1) * 100
diff <- c(baseline, score)
names(diff) <- c("baseline", "score")
diff
```

SVM on PCA without filling missing data with radial kernel
```{r}
model <- svm(y~., data = data.frame(tr_pca[, 1:10], y=tr_y), kernel = "radial", cost = 5, scale = F)
preds_on_CV <- as.numeric(predict(model, newdata = CV_pca[, 1:10]))
score <- rmse(CV_y, preds_on_CV)
(baseline/score - 1) * 100
diff <- c(baseline, score)
names(diff) <- c("baseline", "score")
diff
```


SVM on PCA with filling missing data with radial kernel
```{r}
model <- svm(y~., data = data.frame(tr_pr_pca[, 1:10], y=tr_y), kernel = "radial", cost = 5, scale = F)
preds_on_CV <- as.numeric(predict(model, newdata = CV_pr_pca[, 1:10]))
score <- rmse(CV_y, preds_on_CV)
(baseline/score - 1) * 100
diff <- c(baseline, score)
names(diff) <- c("baseline", "score")
diff
```